---
title: "1.1 knoll stat and plots"
output: html_document
date: "2022-09-21"
---


```{r}
library(tidyverse)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(plotrix)
```

```{r}
load("Data/fish_full_data.Rdata")
load("Data/knoll_mdata.Rdata")
load("Data/coral_cover.Rdata")
load("Data/coral_comm.Rdata")
```

saving paths

```{r}
coral_plots<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R script for report\\Coral plots"


```



# Corals

## plot of mean coral cover + Anova + Tuckey

plot 9 in the report - mean coral cover(soft +stone) for each site


```{r}

# keep relevant column

yearly_cover <- coral_cover %>%
  select(
    Site,
    Year,
    storm,
    Surface_Area,
    Knoll,
    Stony.Coral,
    Hydrozoa,
    Soft.Coral,
    Algae,
    Zoantharia) %>%
  mutate(all_stone = Stony.Coral + Hydrozoa) %>%
  mutate(all_soft = Soft.Coral + Zoantharia)

# calculates mean and se

cover_bar <-
  yearly_cover %>% group_by(Site, Year) %>% 
  summarise(
    "mean_stone" = mean(all_stone, na.rm = T),
    "mean_soft" = mean(all_soft, na.rm = T),
    "n" = n(),
    "SE_stone" = std.error(all_stone),
    "SE_soft" = std.error(all_soft))


cover_bar<-gather(cover_bar,"coral_type","cover",3:4)

cover_bar$Year<-as.factor(cover_bar$Year)

# set the position for se

cover_bar <- cover_bar %>% group_by(Site, Year) %>%
  mutate(
    "se_positon" = case_when(
      coral_type == "mean_stone" ~ cover + SE_stone,
      coral_type == "mean_soft" ~ cumsum(cover) +
        SE_soft
    ))

cover_bar <- cover_bar %>% group_by(Site, Year) %>%
  mutate(
    "se_positon_min" = case_when(
      coral_type == "mean_stone" ~ cover ,
      coral_type == "mean_soft" ~ cumsum(cover)
    )
  )


cover_bar<-cover_bar %>% mutate("site_code"= case_when(
      Site == "Caves" ~ "CA",
      Site == "Princess" ~ "PR",
      Site == "Katza" ~ "KA",
      Site =="NR" ~"NR")
  
)

# plot

ggplot(data = cover_bar,aes(x=Year,y = cover, fill = coral_type))+
    geom_bar(position="stack", stat="identity",color ="black")+
  geom_errorbar(aes(ymax = se_positon,ymin = se_positon_min),width=.1)+
    facet_wrap(site_code~.)+
    theme_bw()+ylab("Mean cover(%)")+ 
  geom_text(data = cover_bar,
        aes(label = n, y = 5),size=6)+
  scale_fill_manual(values = c("#fed0bb","#d77a7d"),name = "Coral type",labels= c("Soft coral","Stony coral"))+
    theme(title = element_text(size = 20),
    axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
    strip.text = element_text(size=22),
    strip.background =element_rect(fill="#faf5ef"))
  
#ggsave("coral_cover_stacked_bar.pdf" , width = 10, height = 8, dpi = 300,path = coral_plots)


```


## model of stony coral cover (%)

before and after the storm,
curate data for coral cover model, omit 2017

those are the chosen models but all models can be found in script 12.2 


```{r}

coral_cover<-coral_cover %>% drop_na(Site)

coral_cover$storm<-ordered(coral_cover$storm,levels = c("Before","After"))

cover_no17<-coral_cover %>% filter(Year != "2017") %>% 
  mutate("knoll_number" = paste(Knoll,Site,sep="_")) %>% na.omit(cover_no17)

cover_no17$Site<-as.factor(cover_no17$Site)

cover_no17$Stony.Coral_p<- cover_no17$Stony.Coral/100
```

### simple model

a simple model that take into account storm,site and knoll id

```{r}
simple_chosen_m <-glmmTMB::glmmTMB(formula = Stony.Coral ~ storm + (1|Site) + (1|knoll_number) , data = cover_no17,
                                  REML = F)

summary(simple_chosen_m)
```

### simple model with site

```{r}
by_site_cover_model<-glmmTMB::glmmTMB(formula = Stony.Coral ~  storm*Site +Max_depth  + (1|knoll_number) ,data = cover_no17,REML = T)


summary(by_site_cover_model)
```

### models for coral richness vs Surface area 

the chosen model, full model selection process is in script 13.1
```{r}

# data prep
only_coral<- coral_comm %>%
  filter(Category %in% c("Soft Coral","Stony Coral","Corallimorpharia","Zoantharia","Hydrozoa")) %>% filter(year != 2017)

only_stone<-only_coral %>% mutate(knoll_number = paste(site,knoll,sep= "_")) %>% filter(knoll_number!= "NR_17",knoll_number!= "NR_11",knoll_number!= "Katza_19") %>% filter(year != 2017) %>%  filter(Category %in% c("Stony Coral","Hydrozoa"))

only_stone<- only_stone %>% group_by(site,knoll,year,storm,Max_depth,Min_depth,Surface_Area) %>% summarise(Species_number = n_distinct(species)) %>%  ungroup() 

# model


model_stone<-lm(formula = log(Species_number) ~ log(Surface_Area)*site +year + Max_depth ,data = only_stone)

summary(model_stone)


# plot

surface_area = rep(seq(from = 1,to = 60,by = 1),16) # Length = 45
site = rep(unique(only_stone$site),each = 240) # Length = 4
year = rep(rep(unique(only_stone$year),each = 60),4) # Length = 4

predict<-data.frame(Site = site,
                       Year = year,
                       Surface_Area= surface_area,
                       Max_depth = rep(4.6))


predict$Surface_Area_log<-log(predict$Surface_Area)

predict_species<-predict(model_stone,predict,type="response")

predict<-cbind(predict,predict_species)

predict$species_number<-exp(predict$predict_species)

predict$Year<- as.factor(predict$Year)

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")

predict <- predict %>% mutate("site_code"= case_when(
      Site == "Caves" ~ "CA",
      Site == "Princess" ~ "PR",
      Site == "Katza" ~ "KA",
      Site =="NR" ~"NR"))

only_stone <- only_stone %>% mutate("site_code"= case_when(
      site == "Caves" ~ "CA",
      site == "Princess" ~ "PR",
      site == "Katza" ~ "KA",
      site =="NR" ~"NR"))

# log scale

ggplot(data = predict,aes(x= Surface_Area,y = species_number,color= site_code))+scale_x_continuous(trans='log2')+scale_y_continuous(trans='log2')+
  theme_classic()+
  geom_smooth(size=1.2,se = T)+
  geom_point(data = only_stone,aes(Surface_Area,Species_number,color = site_code),alpha = 0.3)+
  ylab("Stone coral species")+
  xlab(bquote('Knoll surface area '~(m^2)))+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
  scale_color_manual(name = "Site",values = color)+ guides(color=guide_legend(override.aes=list(fill=NA)))

#ggsave("stone_coral_sp_area.pdf" , width = 14, height = 5, dpi = 300,path = coral_plots)

#ggsave("stone_coral_sp_area.png" , width = 14, height = 5, dpi = 300,path = coral_plots)


```

## Coral rarefactions

#### stone coral by site

richness sbr (by site)

```{r}

# data prep

stone_coral<- coral_comm %>%
  filter(Category %in% c("Stony Coral","Hydrozoa"))


rare_coral<-stone_coral %>% group_by(site,knoll,year,storm,species) %>% summarise("abundance"=sum(points))

data<-spread(rare_coral,species,abundance,fill = 0)


data <- data %>% filter(year != "2017")

data<-data %>% mutate("site"= case_when(
      site == "Caves" ~ "CA",
      site == "Princess" ~ "PR",
      site == "Katza" ~ "KA",
      site =="NR" ~"NR")
  
)
# sbr rarefaction

srb_list<-list()
loop<-1

for (i in unique(data$site)) {
  
  site<- data %>% filter(site == i)
  
  site_richness<-Rarefy::rare_alpha(site[5:ncol(site)],
                          method="hill",
                          q=0,
                          random=200,
                          mean = F)
  
  site_richness$site<-rep(i)
  site_richness$sample<-1:nrow(site_richness)
  
  
  srb_list[[loop]]<-site_richness
  loop<-loop+1
  
  
}

site_richness<-bind_rows(srb_list)

# plot

ggplot(site_richness,aes(x= sample,y=Rarefaction,color=site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Stony coral")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color,name = "Site")+ guides(color=guide_legend(override.aes=list(fill=NA)))

#ggsave("hard_coral_sbr.png" , width = 14, height = 5, dpi = 300,path = coral_plots)

```

shannon sbr (by site)

```{r}
srb_shn_list<-list()
loop<-1

for (i in unique(data$site)) {
  
  site<- data %>% filter(site == i)
  
  site_shanon<-Rarefy::rare_alpha(site[5:ncol(site)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  site_shanon$site<-rep(i)
  site_shanon$sample<-1:nrow(site_shanon)
  
  
  srb_shn_list[[loop]]<-site_shanon
  loop<-loop+1
  
  
}

site_shanon<-bind_rows(srb_shn_list)


ggplot(site_shanon,aes(x= sample,y=Rarefaction,color=site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Stony corals")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color,name = "Site")+ guides(color=guide_legend(override.aes=list(fill=NA)))


#ggsave("hard_coral_shanon_sbr.pdf" , width = 14, height = 5, dpi = 300,path = coral_plots)
```

#### soft coral by site

richness sbr

```{r}
soft_coral<- coral_comm %>%
  filter(Category %in% c("Soft Coral","Zoantharia"))


rare_soft<-soft_coral %>% group_by(site,knoll,year,storm,species) %>% summarise("abundance"=sum(points))

data_soft<-spread(rare_soft,species,abundance,fill = 0)


data_soft <- data_soft %>% filter(year != "2017")

data_soft<-data_soft %>% mutate("site"= case_when(
      site == "Caves" ~ "CA",
      site == "Princess" ~ "PR",
      site == "Katza" ~ "KA",
      site =="NR" ~"NR"))

srb_soft_list<-list()
loop<-1

for (i in unique(data_soft$site)) {
  
  site<- data_soft %>% filter(site == i)
  
  site_richness<-Rarefy::rare_alpha(site[5:ncol(site)],
                          method="hill",
                          q=0,
                          random=200,
                          mean = F)
  
  site_richness$site<-rep(i)
  site_richness$sample<-1:nrow(site_richness)
  
  
  srb_soft_list[[loop]]<-site_richness
  loop<-loop+1
  
  
}

site_soft_richness <- bind_rows(srb_soft_list)


ggplot(site_soft_richness,aes(x= sample,y=Rarefaction,color=site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Soft coral")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color,name = "Site")+
  guides(color=guide_legend(override.aes=list(fill=NA)))


#ggsave("soft_coral_sbr.pdf" , width = 14, height = 5, dpi = 300,path = coral_plots)

```

shannon sbr

```{r}
srb_shn_list_soft<-list()
loop<-1

for (i in unique(data_soft$site)) {
  
  site<- data_soft %>% filter(site == i)
  
  site_shanon<-Rarefy::rare_alpha(site[5:ncol(site)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  site_shanon$site<-rep(i)
  site_shanon$sample<-1:nrow(site_shanon)
  
  
  srb_shn_list_soft[[loop]]<-site_shanon
  loop<-loop+1
  
  
}

site_shanon_soft<-bind_rows(srb_shn_list_soft)


ggplot(site_shanon_soft,aes(x= sample,y=Rarefaction,color=site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Soft corals")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color,name = "Site")+
  guides(color=guide_legend(override.aes=list(fill=NA)))


#ggsave("soft_coral_shanon_sbr.pdf" , width = 14, height = 5, dpi = 300,path = coral_plots)

```


#### stone coral by year

richness sbr

```{r}

l_type<-c("solid","solid","solid","twodash")

srb_list_year<-list()
loop<-1
all_new<-c("#98c1d9","#6b8ead","#003049","#c1121f")


for (i in unique(data$year)) {
  
  year<- data %>% filter(year == i)
  
  year_richness<-Rarefy::rare_alpha(year[5:ncol(year)],
                          method="hill",
                          q=0,
                          random=200,
                          mean = F)
  
  year_richness$Year<-rep(i)
  year_richness$sample<-1:nrow(year_richness)
  
  
  srb_list_year[[loop]]<-year_richness
  loop<-loop+1
  
  
}

year_richness<-bind_rows(srb_list_year)

year_richness$Year<-as.factor(year_richness$Year)

ggplot(year_richness,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year, linetype = Year), size = 1.2) +
  scale_linetype_manual(name = "Year",values = l_type) +
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Stony coral")+
  ylab("Species richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = all_new)+
  guides(color=guide_legend(override.aes=list(fill=NA)))

#ggsave("hard_coral_sbr_year.pdf" , width = 14, height = 5, dpi = 300,path = coral_plots)
```

stone coral shannon by year

```{r}
srb_shanon_list_year<-list()
loop<-1
all_new<-c("#98c1d9","#6b8ead","#003049","#c1121f")


for (i in unique(data$year)) {
  
  year<- data %>% filter(year == i)
  
  year_richness<-Rarefy::rare_alpha(year[5:ncol(year)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  year_richness$Year<-rep(i)
  year_richness$sample<-1:nrow(year_richness)
  
  
  srb_shanon_list_year[[loop]]<-year_richness
  loop<-loop+1
  
  
}

year_richness_sh<-bind_rows(srb_shanon_list_year)

year_richness_sh$Year<-as.factor(year_richness_sh$Year)

ggplot(year_richness_sh,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year, linetype = Year), size = 1.2) +
  scale_linetype_manual(name = "Year",values = l_type)+
  geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Stony coral")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = all_new)+
  guides(color=guide_legend(override.aes=list(fill=NA)))

#ggsave("hard_coral_sbr_year_shanon.pdf" , width = 14, height = 5, dpi = 300,path = coral_plots)
```


