---
title: "2.1 fish analyses for report"
output: html_document
date: "2022-10-02"
---



```{r}
library(tidyverse)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(plotrix)
library(lubridate)
```

```{r}

load("Data/fish_full_data.Rdata")
load("Data/knoll_mdata.Rdata")
load("Data/coral_cover.Rdata")
load("Data/coral_comm.Rdata")

```

```{r}
fish_plots<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R script for report\\Fish plots"
```


# Fish 

## Density 

this is from script 15.1

color scheme

```{r}
#site
color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")
#all years
fish_all<-c("#669bbc","#336683","#003049","#ed3140","#c1121f","#710912")
#all years for katza
fish_k<-c("#669bbc","#336683","#1a4b66","#003049","#ed3140","#c1121f","#710912")

```

1) set one observer

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```



2) filter the north beach

```{r}
one_observer<-one_observer %>%
  filter(Site !="North beach",
         ! Knoll %in% c("Tamar","South bridge reserve","south bridge reserve",
                        "North bridge reserve","north bridge reserve"))
                                      
```

3) add the month to the sample

```{r include=FALSE}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```

### by site

#### fish without schools
bar plot 
```{r}


sfish_2015 <-c("Pseudanthias spp.",
               "Chromis viridis",
               "Neopomacentrus miryae")

ind_per_m<- one_observer %>%
  filter(!Species_2015 %in% sfish_2015) %>% group_by(Site,year_month,Knoll,Surface_Area) %>% summarise("inds"=sum(Abundance)) %>% mutate("inds_m"=inds/Surface_Area)


ind_per_m<-ind_per_m %>% 
  group_by(Site) %>% 
  summarise("mean_density"=mean(inds_m),
            "se"=std.error(inds_m),
             n=n())


ind_per_m <- ind_per_m %>% mutate("Site"= case_when(
      Site == "Caves" ~ "CA",
      Site == "Princess" ~ "PR",
      Site == "Katza" ~ "KA",
      Site =="NR" ~"NR"))

ggplot(data = ind_per_m,aes(x=Site,y=mean_density,fill = Site))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=mean_density-se,ymax=mean_density+se),width =0.1)+
  scale_fill_manual(values = color)+
  geom_text(aes(label = n,y=0.5), size = 7)+  
  xlab("Site")+ ggtitle("Fish abundance (no schools)")+
  ylab(bquote('Mean fish density '(m^2)))+
  theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))

#ggsave("fish_density_m2.pdf" , width = 10, height = 8, dpi = 300,path = fish_plots)
  
```


#### schooling fish 

compare between sites

```{r}

schools_per_m<- one_observer %>%
 filter(Species_2015 %in% sfish_2015) %>%       group_by(Site,year_month,Knoll,Surface_Area) %>%  summarise("inds"=sum(Abundance)) %>%  mutate("inds_m"=inds/Surface_Area)


schools_per_m<-schools_per_m %>%
  group_by(Site) %>% summarise("mean_density"=mean(inds_m),
                               "se"=std.error(inds_m),
                                n=n())

schools_per_m <- schools_per_m %>% mutate("Site"= case_when(
      Site == "Caves" ~ "CA",
      Site == "Princess" ~ "PR",
      Site == "Katza" ~ "KA",
      Site =="NR" ~"NR"))


ggplot(data = schools_per_m,aes(x=Site,y=mean_density,fill = Site))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=mean_density-se,ymax=mean_density+se),width =0.1)+
  scale_fill_manual(values = color)+
  geom_text(aes(label = n,y=0.9), size = 7)+  
  xlab("Site")+ ggtitle("Fish abundance ( schools)")+
  ylab(bquote('Mean schools density '(m^2)))+
  theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))
  
# ggsave("school_density_m2.pdf" , width = 10, height = 8, dpi = 300,path = fish_plots)

```

### by year

#### fish without schools

```{r}
ind_per_m_year<- one_observer %>% filter(!Species_2015 %in% sfish_2015) %>% filter(year_month!="2017 b")%>%  group_by(Site,year_month,Knoll,Surface_Area) %>% summarise("inds"=sum(Abundance)) %>% mutate("inds_m"=inds/Surface_Area)


ind_per_m_year<-ind_per_m_year %>% group_by(year_month) %>% summarise("mean_density"=mean(inds_m),"se"=std.error(inds_m),n=n())

ind_per_m_year$year_month<-as.factor(ind_per_m_year$year_month)

ggplot(data = ind_per_m_year,aes(x=year_month,y=mean_density,fill = year_month))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=mean_density-se,ymax=mean_density+se),width =0.1)+
  geom_text(aes(label = n,y=0.5), size = 7,color= "white")+  
  xlab("Sample")+ ggtitle("Fish abundance (no schools)")+
  ylab(bquote('Mean fish density '(m^2)))+
  theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
  scale_fill_manual(values= fish_all,
                           name = "Sample",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))+
  scale_x_discrete(labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))
  
#ggsave("fish_density_m2_year.png" , width = 10, height = 8, dpi = 300,path = fish_plots)
```

#### total abundance


```{r}
after_storm_knoll<- one_observer %>% filter(year_month =="2020 a") %>% distinct(knoll_id)

after_storm_data_ns<-one_observer %>%
  filter(knoll_id %in% after_storm_knoll$knoll_id) %>%
  filter(!Species_2015 %in% sfish_2015) %>% 
  group_by(Site,year_month) %>%
  summarise("inds" = sum(Abundance))

after_storm_data_ns<-after_storm_data_ns %>% mutate("Site"= case_when(
      Site == "Caves" ~ "CA",
      Site == "Princess" ~ "PR",
      Site == "Katza" ~ "KA",
      Site =="NR" ~"NR"))



ggplot(data = after_storm_data_ns,aes(x=Site,y=inds,color=year_month))+geom_point(size = 3.5)+ggtitle("Fish abundance (no schools)")+
  ylab("Number of individuals")+ theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values= fish_k,
                           name = "Sample",
                           labels = c("Fall 2015", "Fall 2016","Fall 2017",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))


#ggsave("only_storm_knoll_ind_fish_no_schools.png" , width = 10, height = 8, dpi = 300,path =fish_plots)


```



## Biomass

this is from script 15.2

### add biomass column


```{r}
one_observer$class <- with(one_observer,(ceiling(log(Length,2))))


one_observer<-one_observer %>%
  mutate(lenght_class =(2^class)) %>% relocate(lenght_class,.after = 12)

one_observer$midpoint <- ((2^one_observer$class) + (2^(one_observer$class-1)))/2

one_observer<-one_observer %>% relocate(midpoint,.after = 13) %>% relocate(class,.after = 14)

one_observer<- one_observer %>% mutate("true_weight_ind" =  a*Length^b,.after = 14) # to use only when compering years

one_observer<- one_observer %>% mutate("class_weight_ind" =  a*midpoint^b,.after = 15)# to use only when compering sites


# the weight for all ind in the same line

one_observer<-one_observer %>%
  mutate("true_weight_total"=true_weight_ind*Abundance,
         "class_weight_total" = class_weight_ind*Abundance)


one_observer_no_17<-one_observer %>% filter(Year != "2017")


```

### mean biomass by site
no schools

```{r}
biomass_m<-one_observer_no_17 %>%filter(Year !="2015",Year !="2016") %>% filter(!Species_2015 %in% sfish_2015) %>% 
  group_by(Site,year_month,Knoll,Surface_Area) %>% 
  summarise("weight_knoll"=sum(true_weight_total,na.rm = T)) %>% mutate("biomass_m"=weight_knoll/Surface_Area)


biomass_m_site<-biomass_m %>% group_by(Site) %>% summarise("mean_biomass"=mean(biomass_m),
                                                           "se"=std.error(biomass_m),
                                                           "n"=n())

biomass_m_site <-biomass_m_site %>% mutate("Site"= case_when(
      Site == "Caves" ~ "CA",
      Site == "Princess" ~ "PR",
      Site == "Katza" ~ "KA",
      Site =="NR" ~"NR"))

ggplot(data = biomass_m_site,aes(x=Site,y=mean_biomass,fill = Site))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=mean_biomass-se,ymax=mean_biomass+se),width =0.1)+
  scale_fill_manual(values = color)+
  geom_text(aes(label = n,y=20), size = 7)+  
  xlab("Site")+ ggtitle("Fish biomass (no schools)")+
  ylab(bquote('Mean fish biomass '(g/m^2)))+
  theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))

#ggsave("fish_biomass_per_m_site.png" , width = 10, height = 8, dpi = 300,path = fish_plots)

```

### mean biomass by year
no schools

```{r}
biomass_m_year<-biomass_m %>% group_by(year_month) %>% summarise("mean_biomass"=mean(biomass_m),                               "se"=std.error(biomass_m),                                    "n"=n())


ggplot(data = biomass_m_year,aes(x=year_month,y=mean_biomass,fill = year_month))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=mean_biomass-se,ymax=mean_biomass+se),width =0.1)+
  geom_text(aes(label = n,y=20), size = 7,color= "white")+  
  xlab("Sample")+ ggtitle("Fish biomass (no schools)")+
  ylab(bquote('Mean fish biomass '(g/m^2)))+
  theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
  scale_fill_manual(values= fish_all[3:6],
                           name = "Sample",
                           labels = c("Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))+
  scale_x_discrete(labels = c("Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))
#ggsave("fish_biomass_per_m_year.png" , width = 10, height = 8, dpi = 300,path = fish_plots)
```



###biomass by lenght density plot

```{r}
after_storm_knoll<- one_observer_no_17 %>%
  filter(year_month =="2020 a") %>%
  distinct(knoll_id)

after_knolls<-one_observer_no_17 %>%
  filter(knoll_id %in% after_storm_knoll$knoll_id) %>% filter(year_month != "2015 b",year_month != "2016 b") %>% filter(!Species_2015 %in% sfish_2015 )

ggplot(after_knolls, aes(x= log(true_weight_total/1000), fill=year_month)) + geom_density(alpha =0.8)+
  ggtitle("Fish biomass")+ xlab("Log(Weight (kg))")+
  ylab("Density")+ theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_fill_manual(values= fish_all[3:7],
                           name = "Sample",
                           labels = c("Fall 2018",
                                      "Spring 2020",
                                       "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("fish_biomass_dist_year.png" , width = 10, height = 8, dpi = 300,path = fish_plots)

after_knolls_res<-after_knolls %>%  filter(Status !="T")

ggplot(after_knolls_res, aes(x= log(true_weight_total/1000), fill=year_month)) + geom_density(alpha =0.8)+
  ggtitle("Fish biomass")+ xlab("Log(Weight (kg))")+
  ylab("Density")+ theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_fill_manual(values= fish_all[3:7],
                           name = "Sample",
                           labels = c("Fall 2018",
                                      "Spring 2020",
                                       "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("fish_biomass_dist_year_only_res.png" , width = 10, height = 8, dpi = 300,path = fish_plots)
```

### total biomass by year

```{r}
total_biomass_storm <- after_knolls %>%
  filter(!Species_2015 %in% sfish_2015) %>%  group_by(Site,year_month) %>%  summarise("total_biomass"=sum(true_weight_total,na.rm = T)) %>% filter(year_month != "2015 b",year_month != "2016 b") 


total_biomass_storm<-total_biomass_storm %>% mutate("Site"= case_when(
      Site == "Caves" ~ "CA",
      Site == "Princess" ~ "PR",
      Site == "Katza" ~ "KA",
      Site =="NR" ~"NR")
  
  
)


ggplot(data = total_biomass_storm,aes(x=Site,y=total_biomass/1000,color=year_month))+geom_point(size = 7,alpha = 0.8)+ggtitle("Fish biomass")+
  ylab("Total biomass (kg)")+ theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values= fish_all[3:7],
                           name = "Sample",
                           labels = c("Fall 2018",
                                      "Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))

#ggsave("fish_biomass_point_year.png" , width = 10, height = 8, dpi = 300,path = fish_plots)

```


## Species area relationship

I cant find the original script for the plots but i will complie it now based on the model describe in the ms
 
 ill start by cleaning the environment and create the data for the sp-area plots and models
 
```{r}
rm(list=ls())

load("Data/fish_full_data.Rdata")

fish_plots<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R script for report\\Fish plots"

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")
#all years
fish_all<-c("#669bbc","#336683","#003049","#ed3140","#c1121f","#710912")
#all years for katza
fish_k<-c("#669bbc","#336683","#1a4b66","#003049","#ed3140","#c1121f","#710912")

```

```{r}

# remove confidence 1

sp_area<-fish_full_data
sp_area$Confidence[is.na(sp_area$Confidence)]<-(0)
sp_area<- sp_area %>% filter(Confidence < 1)


# remove irrgular sites


knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

sp_area<-sp_area %>% filter(!Knoll %in% knolls_to_remove) %>% filter(Site !="North beach")

# choose one observer


one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(sp_area$survey_id)) {
  
  data <- sp_area %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)


# keep only residence fish

only_cryptic <- one_observer %>% filter(Status != "T")

# calculate speciess richness per knoll

species_knoll<- only_cryptic %>% group_by(survey_id) %>% mutate(Species_number = n_distinct(Species_2015)) %>% select(-Species,-Species_2015,-Family,-Abundance,-Length,-Confidence,-Status,-a,-b) %>%  ungroup() %>% distinct(.keep_all = T)

# remove 2017 as only katza was sampled

species_knoll<-species_knoll %>% filter(Year != 2017)

# add the month, season and convert to factors

species_knoll$month<-month(species_knoll$Date)
species_knoll$month<-as.factor(species_knoll$month)


species_knoll$Season <- ifelse(species_knoll$month == 6,"Spring","Fall")

species_knoll$Year<-as.factor(species_knoll$Year)
species_knoll$Site<-as.factor(species_knoll$Site)
species_knoll$storm<-as.factor(species_knoll$storm)
species_knoll$year_month<-as.factor(species_knoll$year_month)


# remove outliars knolls

no_nr_outliar<-species_knoll %>% filter(knoll_id != "Katza_19",knoll_id != "NR_17")


```

model as described in the report draft

```{r}

model<-lm(formula = log(Species_number) ~ log(Surface_Area)*Site + year_month + Max_depth ,data = no_nr_outliar)



```

plots

```{r}

# plot by site

model_gg_site=ggeffects::ggpredict(model,terms = c('Surface_Area[1:66 by=1]','Site'))

ggplot()+
  theme_classic()+
  geom_line(data = model_gg_site,aes(x= x,y = predicted,color= group),size=1.5)+ 
  geom_point(data = species_knoll,aes(Surface_Area,Species_number,color = Site),alpha = 0.5)+scale_x_continuous(trans='log2')+scale_y_continuous(trans='log2')+
  ylab("Richness")+
  xlab(bquote('Knoll surface area '~(m^2)))+
    theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
  scale_color_manual(name = "Site",values = color,
                     labels = c("CA", "KA",
                                      "NR","PR"))
# ggsave("fish_sp_area_residence_site.png" , width = 14, height = 5, dpi = 300,path = fish_plots)

  
# plot by year

model_gg_year=ggeffects::ggpredict(model,terms = c('Surface_Area[1:66 by=1]','year_month'))

ggplot()+
  theme_classic()+
  geom_line(data = model_gg_year,aes(x= x,y = predicted,color= group),size=1.5)+ 
  geom_point(data = species_knoll,aes(Surface_Area,Species_number,color = year_month),alpha = 0.5)+
  scale_x_continuous(trans='log2')+scale_y_continuous(trans='log2')+
  ylab("Richness")+
  xlab(bquote('Knoll surface area '~(m^2)))+
    theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values=fish_all,
                           name = "Sample",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))+
        scale_fill_manual(values=fish_all,
                           name = "Sample",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))

# ggsave("fish_sp_area_residence_year.png" , width = 14, height = 5, dpi = 300,path = fish_plots)

```




*Note* - this sp-area by year plot is different from the one presented in the paper because there is no interaction between surface area and year_month. the difference is that in the current plot the storm curve remain always below the others while when modeling the interaction we see the posts1 convegre with the rest... 

## Rarefactions


```{r}

rm(list=ls())

load("Data/fish_full_data.Rdata")

fish_plots<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R script for report\\Fish plots"

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")
#all years
fish_all<-c("#669bbc","#336683","#003049","#ed3140","#c1121f","#710912")
#all years for katza
fish_k<-c("#669bbc","#336683","#1a4b66","#003049","#ed3140","#c1121f","#710912")

```

prepare the data

```{r}

# choose one observer

one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)

#filter low confidance


one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <1)


#filter irregular sites

one_observer<-one_observer %>%
  filter(Site !="North beach",
         ! Knoll %in% c("Tamar","South bridge reserve","south bridge reserve",
                        "North bridge reserve","north bridge reserve"))

# add month

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


# create wide data for 2015 resolution

all_fish_2015 <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_2015<-spread(all_fish_2015,Species_2015,aboundance,fill = 0)


```



#### by site

##### Richness

```{r}
srb_list<-list()
loop<-1

wide_2015<-wide_2015 %>% filter(Year !="2017")

for (i in unique(wide_2015$Site)) {
  
  site<- wide_2015 %>% filter(Site == i)
  
  site_richness<-Rarefy::rare_alpha(site[9:ncol(site)],
                          method="hill",
                          q=0,
                          random=200,
                          mean = F)
  
  site_richness$Site<-rep(i)
  site_richness$sample<-1:nrow(site_richness)
  
  
  srb_list[[loop]]<-site_richness
  loop<-loop+1
  
  
}

site_richness<-bind_rows(srb_list)


ggplot(site_richness,aes(x= sample,y=Rarefaction,color=Site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color,labels = c("CA", "KA","NR","PR"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))


# ggsave("fish_sbr.pdf" , width = 14, height = 5, dpi = 300,path = fish_plots)
```

##### Shannon with schools

```{r}
srb_list_shannon<-list()
loop<-1

wide_2015<-wide_2015 %>% filter(Year !="2017")

for (i in unique(wide_2015$Site)) {
  
  site<- wide_2015 %>% filter(Site == i)
  
  site_shannon<-Rarefy::rare_alpha(site[9:ncol(site)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  site_shannon$Site<-rep(i)
  site_shannon$sample<-1:nrow(site_shannon)
  
  
  srb_list_shannon[[loop]]<-site_shannon
  loop<-loop+1
  
  
}

site_richness_shannon<-bind_rows(srb_list_shannon)



ggplot(site_richness_shannon,aes(x= sample,y=Rarefaction,color=Site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color,labels = c("CA", "KA","NR","PR"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))



#ggsave("fish_shannon_sbr.png" , width = 14, height = 5, dpi = 300,path = fish_plots)
```
##### Shannon no schools

```{r}

srb_list_shannon_ns<-list()
loop<-1

wide_2015_ns<-wide_2015 %>% filter(Year !="2017")

wide_2015_ns$`Pseudanthias spp.`<-NULL
wide_2015_ns$`Chromis viridis`<-NULL
wide_2015_ns$`Neopomacentrus miryae`<-NULL

for (i in unique(wide_2015_ns$Site)) {
  
  site<- wide_2015_ns %>% filter(Site == i)
  
  site_shannon_ns<-Rarefy::rare_alpha(site[9:ncol(site)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  site_shannon_ns$Site<-rep(i)
  site_shannon_ns$sample<-1:nrow(site_shannon_ns)
  
  
  srb_list_shannon_ns[[loop]]<-site_shannon_ns
  loop<-loop+1
  
  
}

site_richness_shannon_ns<-bind_rows(srb_list_shannon_ns)


ggplot(site_richness_shannon_ns,aes(x= sample,y=Rarefaction,color=Site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.05, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish - no schools")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color,labels = c("CA", "KA","NR","PR"))+
  guides(color=guide_legend(override.aes=list(fill=NA)))


#ggsave("fish_shannon_sbr_no_schools.png", width = 14, height = 5, dpi = 300,path = fish_plots)
```


#### By year

##### Richness

```{r}
l_type<-c("solid","solid","solid","twodash","solid","solid")

srb_list_year<-list()
loop<-1

wide_2015<-wide_2015 %>% filter(Year !="2017")

for (i in unique(wide_2015$year_month)) {
  
  year<- wide_2015 %>% filter(year_month == i)
  
  year_richness<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=0,
                          random=200,
                          mean = F)
  
  year_richness$Year<-rep(i)
  year_richness$sample<-1:nrow(year_richness)
  
  
  srb_list_year[[loop]]<-year_richness
  loop<-loop+1
  
  
}

year_richness<-bind_rows(srb_list_year)

labs(color  = "Year", linetype = "Year")


ggplot(year_richness,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year, linetype = Year), size = 1.2) +
  scale_linetype_manual(name = "Sample",values = l_type,labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021")) +
  geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Richness")+ 
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = fish_all,name = "Sample",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))+

  guides(color=guide_legend(override.aes=list(fill=NA)))


#ggsave("fish_sbr_year.pdf" , width = 14, height = 5, dpi = 300,path = fish_plots)
```

##### Shannon with schools

```{r}
srb_list_year_sh<-list()
loop<-1

wide_2015<-wide_2015 %>% filter(Year !="2017")

for (i in unique(wide_2015$year_month)) {
  
  year<- wide_2015 %>% filter(year_month == i)
  
  year_richness_shannon<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  year_richness_shannon$Year<-rep(i)
  year_richness_shannon$sample<-1:nrow(year_richness_shannon)
  
  
  srb_list_year_sh[[loop]]<-year_richness_shannon
  loop<-loop+1
  
  
}

year_richness_shannon<-bind_rows(srb_list_year_sh)

labs(color  = "Year", linetype = "Year")

ggplot(year_richness_shannon,
       aes(x = sample, y = Rarefaction, color = Year)) +
  geom_line(aes(color = Year,linetype = Year), size = 1.2) +
  scale_linetype_manual(name = "Sample",
                        values = l_type,
                        labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))+
  geom_ribbon(aes(ymin = IC_low, ymax = IC_up),
              alpha = 0.1,
              linetype = 0) +
  theme_bw() + xlab("Random sample number") + ggtitle("Fish") +
  ylab("Shannon diversity") +
  theme(
    axis.text.x = element_text(size = 20),
    plot.title = element_text(size = 24),
    axis.text.y = element_text(size = 20),
    axis.title.x = element_text(size = 22),
    axis.title.y = element_text(size = 22),
    legend.title = element_text(size = 22),
    legend.text = element_text(size = 20)) + 
    scale_color_manual(values = fish_all,name = "Sample",
    labels = c(
      "Fall 2015",
      "Fall 2016",
      "Fall 2018",
      "Spring 2020",
      "Fall 2020" ,
      "Spring 2021"))+
      guides(color = guide_legend(override.aes = list(fill = NA)))
             

#ggsave("fish_sbr_year_shannon.pdf" , width = 14, height = 5, dpi = 300,path = fish_plots)
```


##### Shannon no schools

```{r}
srb_list_year_sh_nosc<-list()
loop<-1

wide_2015_ns<-wide_2015 

wide_2015_ns$`Pseudanthias spp.`<-NULL
wide_2015_ns$`Chromis viridis`<-NULL
wide_2015_ns$`Neopomacentrus miryae`<-NULL


for (i in unique(wide_2015_ns$year_month)) {
  
  year<- wide_2015_ns %>% filter(year_month == i)
  
  year_richness_shannon_ns<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  year_richness_shannon_ns$Year<-rep(i)
  year_richness_shannon_ns$sample<-1:nrow(year_richness_shannon_ns)
  
  
  srb_list_year_sh_nosc[[loop]]<-year_richness_shannon_ns
  loop<-loop+1
  
  
}

year_richness_shannon_ns<-bind_rows(srb_list_year_sh_nosc)


ggplot(year_richness_shannon_ns,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year,linetype = Year), size = 1.2) +
  scale_linetype_manual(name = "Sample",
                        values = l_type,
                        labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish - without schools")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = fish_all,name = "Sample",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))+
  guides(color = guide_legend(override.aes = list(fill = NA)))
  


#ggsave("fish_sbr_year_shannon_no_s.pdf" , width = 14, height = 5, dpi = 300,path = fish_plots)
```

## nMDS

```{r}
rm(list=ls())

load("Data/fish_full_data.Rdata")

fish_plots<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R script for report\\Fish plots"

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")

new_6<-c("#277da1",
         "#4d194d",
         "#43aa8b",
         "#d9ed92",
         "#f94144",
         "#f8961e")

new_7<-c("#277da1",
         "#4d194d",
         "#90be6d",
         "#43aa8b",
         "#d9ed92",
         "#f94144",
         "#f8961e")

scales::show_col(new_6)
scales::show_col(new_7)

```


```{r}
# one observer

one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)

# confidance

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

# month


one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


#

```




```{r}

cryp_fish_2015 <- one_observer %>% filter(Status != "T") %>% 
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_cryptic_2015<-spread(cryp_fish_2015,Species_2015,aboundance,fill = 0)

wide_cryptic_2015<-wide_cryptic_2015 %>% drop_na(Surface_Area)

wide_cryptic_2015<- wide_cryptic_2015 %>% mutate("Site"= case_when(
      Site == "Caves" ~ "CA",
      Site == "Princess" ~ "PR",
      Site == "Katza" ~ "KA",
      Site =="NR" ~"NR"))


rm(cryp_fish_2015)


wide_cryptic_2015$year_month <-
  as.character(wide_cryptic_2015$year_month)


wide_cryptic_2015$year_month <-
  ifelse(month(wide_cryptic_2015$Date) == 6,
         paste0("Spring ", year(wide_cryptic_2015$Date)),
         paste0("Fall " , year(wide_cryptic_2015$Date)))


wide_cryptic_2015$year_month<-as.factor(wide_cryptic_2015$year_month)


wide_cryptic_2015$year_month <-
  ordered(
    wide_cryptic_2015$year_month,
    levels = c(
      "Fall 2015",
      "Fall 2016",
      "Fall 2017",
      "Fall 2018",
      "Spring 2020",
      "Fall 2020" ,
      "Spring 2021"
    )
  )




```

#### By site

```{r}


# make the plot for all sites execpt Kazta (6 colors) 

wide_no_k<-wide_cryptic_2015 %>% filter(Site!="KA")
wide_no_k$year_month<-droplevels(wide_no_k$year_month)

for (i in unique(wide_no_k$Site)) {
  
  data <- wide_no_k %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  ord <-
    metaMDS(
      wis_data,
      distance = "bray" ,
      trace = F,
      k = 2,
      autotransform = FALSE
    ) 
  
  color_y=new_6
 
  Year <- factor(data$year_month)
  
  Year<-droplevels(Year)

  plot_name<-paste0(fish_plots,"\\",i,"_site",".pdf")

pdf(file = plot_name,   # The directory you want to save the file in
    width = 8,          # The width of the plot in inches
    height = 8)    
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.5,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

dev.off()

}


# make the nmds plot for katza (7 colors)

wide_k<-wide_cryptic_2015 %>% filter(Site=="KA")

for (i in unique(wide_k$Site)) {
  
  data <- wide_k %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE) 
  
  color_y=new_7
 
  Year <- factor(data$year_month)
  
  Year<-droplevels(Year)
  
  plot_name<-paste0(fish_plots,"\\",i,"_site",".pdf")

pdf(file = plot_name,   # The directory you want to save the file in
    width = 8,          # The width of the plot in inches
    height = 8)  
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.5,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

dev.off()

}


```

#### By year

```{r}

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(wide_cryptic_2015$year_month)) {
  
  data <- wide_cryptic_2015 %>% filter(year_month == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  site <- factor(data$Site)
  
plot_name<-paste0(fish_plots,"\\",i,"_year",".pdf")

pdf(file = plot_name,   # The directory you want to save the file in
    width = 7,          # The width of the plot in inches
    height = 7)  
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

dev.off()

}
```

